Supernovae Time Profiles as a Probe of New Physics at Neutrino Telescopes

### 2024-02-21 paper/sections/detector_response.tex
Updates from Overleaf

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/hits_and_likelihood.pdf}
    \caption{\textbf{\textit{Timing profile of hits and test statistic as a function of the timing window.}}
    The hit rate from the SM neutrino flux peaks at around 0.1~s, coinciding with the peak emission of $\bar{\nu}_{e}$ from the SN.
    The shaded regions correspond to the timing windows that give peak sensitivity to the particular model.
    The low-mass models can give more prompt signals since they decay quickly and travel luminally, while the subliminal speeds and relatively long decay times for larger-mass models lead to a delayed signal.
    In the right panel, we show the test statistic as a function of the starting time and length of the window.
    The relatively broad distribution of high test-statistic values shows that this analysis technique does not require extremely precise reconstruction of the $t_{\mathrm{bounce}}$.
    These are the same points in the parameter space as Fig.~\ref{fig:fluxes}.
    }
    \label{fig:hits_and_likelihood}
\end{figure*}
\
\textbf{\textit{Detector Response and Statistical Treatment}}---The IceCube Neutrino Observatory comprises 5,160 light-detecting digital optical modules (DOMs) buried in a cubic kilometer of the deep, transparent Antarctic ice sheet.
When neutrinos interact within or near the detector, the charged by-products emit photons that the DOMs can detect.
The DOMs are arranged on 86 strings of 60 DOMs with an inter-string distance between 70~m and 125~m.
This enables IceCube to resolve individual neutrinos with energies approximately ranging between a few GeV and a few PeV.
The neutrinos produced by SNe are far below this threshold, and the neutrinos cannot be individually resolved; however, the immense number of neutrinos produced in an SN produces an increase in the single-photon rate of the detector.
This increase in the rate of photons can be distinguished by the background caused by dark noise and radioactive activity in the DOM glass to enable the detection of galactic SNe events~\cite{Griswold:2023iwz}.

New neutrino physics will affect the development of an SN and distort the temporal structure of the photon signal seen in the IceCube detector.
Thus, one may look for excess events in certain time intervals as evidence of this new physics.
In this work, we simulate the light curve produced by a standard SN with a progenitor mass of $8.8~M_{\odot}$ and those produced by different BSM scenarios; see the left-hand panel of Fig.~\ref{fig:hits_and_likelihood} for example light curves.
Since more than 93\% of the photons detected in IceCube originate in inverse beta decay interactions, the SM-only event rate peaks at around 0.1~s.
BSM scenarios, on the other hand, can give rise to a prompt or delayed signal depending on the particulars of the model.
For instance, since low-mass majorons have fast decay times, they can escape the SN early and decay to an all-flavor flux of neutrinos, producing $\bar{\nu}_{e}$ before the SM peak.
The higher-mass majorons, on the other hand, decay less quickly and acquire a time decay from traveling sub-luminally, thus producing a later signal.
We use the \texttt{ASTERIA}~\cite{spencer_griswold_2020_3926835} package to simulate the detector response to the different SNe scenarios and the thermal and radioactive noise in the DOMs.
This package simulates light yields from coherent $\nu_{e}$ scattering off electrons in the ice, inverse beta decay of $\bar{\nu}_{e}$ on nuclei in the ice, and charged- and neutral-current interactions for $\nu_{\alpha}$.
In order to interface to this package, we use \texttt{SNEWPY}~\cite{baxter2021snewpy} package and, in particular, the \texttt{ParametrizedFlux} object.
In addition to photons produced in neutrino interactions, this package also simulates photons from thermal noise in the IceCube PMTs and from radioactive decay in the pressure glass housing.

With the number of hits in the detector from noise, SM-only scenarios, and BSM scenarios as a function of time, we can then quantify the probability of seeing a certain number of hits in a given time window with:
\begin{equation}
    -2\Delta\mathrm{LLH} = 2 \left[N_{\mathrm{exp.}} - N_{\mathrm{obs.}} + N_{\mathrm{obs.}}\log\left(\frac{N_{\mathrm{obs.}}}{N_{\mathrm{exp.}}}\right)\right].
    % -2\Delta\mathrm{LLH}(t_{\mathrm{start}}, \Delta t) = 2 \left[N_{\mathrm{exp.}} - N_{\mathrm{obs.}} + N_{\mathrm{obs.}}\log\left(\frac{N_{\mathrm{obs.}}}{N_{\mathrm{exp.}}}\right)\right].
\end{equation}
where $N_{\mathrm{obs.}}$ is the number of photons seen in the detector in the given time given and $N_{\mathrm{exp.}}$ is the number of photons expected from a particular BSM hypothesis.
In the right-hand panel of Fig.~\ref{fig:hits_and_likelihood}, we show the likelihood space as a function of the time at which the time range begins, $t_{\mathrm{start}}$ and the duration of the time range, $\Delta t$ for an example BSM hypothesis.

For each physics hypothesis, we then select the time range that maximizes the test statistic and use that maximal test statistic value.
The optimal ranges for each physics hypothesis are shown by the shaded regions in the left panel of Fig.~\ref{fig:hits_and_likelihood}. Furthermore, the test statistic value relative to the SM-only hypothesis is shown in the right-hand panel of Fig.~\ref{fig:hits_and_likelihood}.
The distribution of test-statistic values is not sharply peaked, and thus, this analysis retains much of its sensitivity even with errors on the reconstruction of $t_{\mathrm{bounce}}$ $\mathcal{O}\left(1~\mathrm{s}\right)$.

>>>>

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/hits_and_likelihood.pdf}
    \caption{\textbf{\textit{Timing profile of hits and test statistic as a function of the timing window.}}
    The hit rate from the SM neutrino flux peaks at around 0.1~s, coinciding with the peak emission of $\bar{\nu}_{e}$ from the SN.
    The shaded regions correspond to the timing windows that give peak sensitivity to the particular model.
    In the right panel, we show the test statistic as a function of the starting time and length of the window.
    The relatively broad distribution of high test-statistic values shows that this analysis technique does not require extremely precise reconstruction of the $t_{\mathrm{bounce}}$.
    These are the same points in the parameter space as Fig.~\ref{fig:fluxes}.
    }
    \label{fig:hits_and_likelihood}
\end{figure*}
\
\textbf{\textit{Detector Response and Statistical Treatment}}---The IceCube Neutrino Observatory comprises 5,160 light-detecting digital optical modules (DOMs) buried in a cubic kilometer of the deep, transparent Antarctic ice sheet.
When neutrinos interact within or near the detector, the charged by-products emit photons that the DOMs can detect.
The DOMs are arranged on 86 strings of 60 DOMs with an inter-string distance between 70~m and 125~m.
This enables IceCube to resolve individual neutrinos with energies approximately ranging between a few GeV and a few PeV.
The neutrinos produced by SNe are far below this threshold, and the neutrinos cannot be individually resolved; however, the immense number of neutrinos produced in an SN increases the single-photon rate of the detector.
This dramatic increase in the rate of photons can be distinguished from the background caused by dark noise and radioactive activity in the DOM glass to enable the detection of galactic CCSNe events~\cite{Griswold:2023iwz}.

New neutrino physics will affect the development of an SN and distort the temporal structure of the photon signal seen in the IceCube detector.
Thus, one may look for excess events in certain time intervals as evidence of this new physics.
In this work, we simulate the light curve produced by a standard SN with a progenitor mass of $8.8~M_{\odot}$ and those produced by different BSM scenarios aforementioned; see the left-hand panel of Fig.~\ref{fig:hits_and_likelihood} for light curves of the same two benchmarks discussed above.
Since more than 93\% of the photons detected in IceCube originate in inverse beta decay interactions, the SM-only event rate peaks at around 0.1~s.
BSM scenarios, on the other hand, can give rise to an early or delayed signal depending on the particulars of the model.
In essence, Majorons can be generated before the peak of SM $\bar{\nu}_{e}$ flux and as light majorons can travel nearly relativistically, they can escape the SN with negligible time-delay then decay to an all-flavor flux of neutrinos, producing early $\bar{\nu}_{e}$ fluxes before the SM case.
The higher-mass majorons, on the other hand acquire a time decay from traveling sub-luminally and at a large detour, thus producing a later signal with sizable time delay.
We use the \texttt{ASTERIA}~\cite{spencer_griswold_2020_3926835} package to simulate the detector response to the different SNe scenarios and the thermal and radioactive noise in the DOMs.
 This package simulates light yields from coherent $\nu_{e}$ scattering off electrons in the ice, inverse beta decay of $\bar{\nu}_{e}$ on nuclei in the ice, and charged- and neutral-current interactions for $\nu_{\alpha}$.
In order to interface to this package, we use \texttt{SNEWPY}~\cite{SNEWS:2021ewj} package and, in particular, the \texttt{ParametrizedFlux} object.
In addition to photons produced in neutrino interactions, this package also simulates photons from thermal noise in the IceCube PMTs and from radioactive decay in the pressure glass housing.

With the number of hits in the detector from noise, SM-only scenarios, and BSM scenarios as a function of time, we can then quantify the probability of seeing a certain number of hits in a given time window with:
\begin{equation}
    -2\Delta\mathrm{LLH} = 2 \left[N_{\mathrm{exp.}} - N_{\mathrm{obs.}} + N_{\mathrm{obs.}}\log\left(\frac{N_{\mathrm{obs.}}}{N_{\mathrm{exp.}}}\right)\right].
    % -2\Delta\mathrm{LLH}(t_{\mathrm{start}}, \Delta t) = 2 \left[N_{\mathrm{exp.}} - N_{\mathrm{obs.}} + N_{\mathrm{obs.}}\log\left(\frac{N_{\mathrm{obs.}}}{N_{\mathrm{exp.}}}\right)\right].
\end{equation}
where $N_{\mathrm{obs.}}$ is the number of photons seen in the detector in the given time and $N_{\mathrm{exp.}}$ is the number of photons expected from a particular BSM hypothesis.
In the right-hand panel of Fig.~\ref{fig:hits_and_likelihood}, we show the likelihood space as a function of the time at which the time range begins, $t_{\mathrm{start}}$ and the duration of the time range, $\Delta t$ for an example BSM hypothesis.

For each physics hypothesis, we then select the time range that maximizes the test statistic and use that maximal test statistic value. The optimal ranges for each physics hypothesis are shown by the shaded regions in the left panel of Fig.~\ref{fig:hits_and_likelihood}, which is around $\unit[0.1]{s}$ ($\unit[10]{s}$) for the light (heavy) benchmark case, consistent with its time spread of $\bar{\nu}_e$ flux produced. Furthermore, the test statistic value relative to the SM-only hypothesis is shown in the right-hand panel of Fig.~\ref{fig:hits_and_likelihood}.
The distribution of test-statistic values is not sharply peaked, and thus, this analysis retains much of its sensitivity even with errors on the reconstruction of $t_{\mathrm{bounce}}\sim\mathcal{O}\left(10^{-3}~\mathrm{s}\right)$ \cite{Halzen_2009}.

### 2024-02-21 paper/sections/introduction.tex
Updates from Overleaf

\textbf{\textit{Introduction}}---
The Standard Model (SM) is a remarkable but incomplete theory. Puzzles such as the non-vanishing tiny neutrino mass, the origin of observed matter-antimatter asymmetry, and the nature of dark matter, among others, all seek explanations beyond the Standard Model (BSM). However, there are no affirmative clues for the energy scale the yet undiscovered particles should appear. Numerous BSM searches across the scales have been performed, from collider searches at high energies \cite{Nath:2010zj} to studies of cosmic microwave background at temperatures near absolute zero \cite{Baumann:2015rya}. Galactic SN explosion, where majority of emitted energy is released in the form of neutrinos, offers unique opportunities in testing BSM physics via interactions with active neutrinos. So far we have observed SN 1987A in neutrinos. The duration of such a signal \cite{Kamiokande-II:1987idp,Bionta:1987qt,Baksan} as well as the inferred total emitted energy \cite{Loredo:2001rx,Pagliaroli:2008ur,Huedepohl2010} allow us already to set a tentative constraint on the energies that could be taken away from any light BSM particle copiously produced in the interior of a star. Indeed, many different BSM scenarios models have been constrained in this way, see \emph{e.g.} \cite{Raffelt:2011nc,Arguelles:2016uwb,Suliga:2020vpz,Lucente:2021hbp,Caputo:2022rca,Caputo:2021rux,PhysRevD.100.083002,DeRocco:2019njg,Kazanas:2014mca,Magill:2018jla}.

If, additionally, new physics particles, emitted from the SN core, can decay to neutrinos, even stronger limits can be set. In particular, for the standard case, neutrinos interact frequently in the core and eventually leave the star with $\mathcal{O}(10)$ MeV energy. New states after being produced can stream out without further interactions. Consequently, they typically exit with higher energies, subsequently decaying into neutrinos with an energy on the order of $\mathcal{O}(100)$ MeV. Limits can be set from the fact that such high energy neutrinos were not recorded from SN 1987A \cite{Kamiokande-II:1987idp,Bionta:1987qt,Baksan, Fiorillo:2022cdq, Brdar:2023tmi}. Once produced from SN, the new state can be non-relativistic and travels slowly, causing time delays to the arrival time of its daughter particle which was also recently employed \cite{Brdar:2023tmi} for a time window around one day. However, as being decoupled immediately, it is possible that new states can exit the supernova at an earlier time instead—an effect that has not been previously considered. These two effects in combination would cause advanced or delayed neutrino signals to the standard case.

On the other hand, neutrino telescopes such as IceCube have not only pioneered in detecting neutrinos at ultra-high energies \cite{IceCube:2014stg,IceCube:2018cha} and from specific extragalactic sources \cite{IceCube:2023ame,IceCube:2022der}, but also in testing temporal correlation of MeV neutrino events with fast radio bursts with its precise timing measurements \cite{IceCube:2019acm}. In this work, we demonstrate that such timing measurements will play a crucial role in tracking the delayed neutrino signals, and interestingly reveil new features of advanced neutrino signals coming from the immediate-decoupling of its mother particles. This novel timing patterns generically exist in models with new weakly interacting particles, such as sterile neutrino via the dipole portal \cite{Magill:2018jla,Brdar:2020quo,Brdar:2023tmi} or mixing portals \cite{Arg_elles_2019, Suliga_2020}, Majoron models \cite{Fiorillo:2022cdq}. Once produced inside the SN, these particles stream out at an earlier time than the standard neutrino signal. Meanwhile, for non-negligible mass, these particle would travel slower than active neutrino before decaying or converting back to active neutrinos. Via the timing measurement, we will show that neutrino telescope can provide unprecedented constraints on the parameter space in BSM.\\

>>>>

\textbf{\textit{Introduction}}---
The Standard Model (SM) is a remarkable but incomplete theory. Puzzles such as the non-vanishing neutrino mass, the origin of observed matter-antimatter asymmetry, and the nature of dark matter, among others, seek explanations beyond the Standard Model (BSM). There are also no firmly established clues on the energy scale at which the new particles should appear. To this end, numerous BSM searches across the scales have been performed, from collider searches at high energies \cite{Nath:2010zj} to studies of cosmic microwave background at temperatures near absolute zero \cite{Baumann:2015rya}. Core collapse supernovae (CCSNe) release most energy in neutrinos, offering unique opportunities to test BSM physics via interactions with neutrinos.
The duration of the burst~\cite{Kamiokande-II:1987idp,Bionta:1987qt,Baksan} and inferred total energy of neutrinos~\cite{Loredo:2001rx,Pagliaroli:2008ur,Huedepohl2010} from SN 1987A---the only CCSN observed in neutrinos---have already given tentative constraints on the energies that could be taken away by any light BSM particle copiously produced in the interior of a star.
Indeed, many different BSM scenarios models have been constrained in this way, see \textit{e.g.}~\cite{Raffelt:2011nc,Arguelles:2016uwb,Suliga:2020vpz,Lucente:2021hbp,Caputo:2022rca,Caputo:2021rux,PhysRevD.100.083002,DeRocco:2019njg,Kazanas:2014mca,Magill:2018jla}.

If, additionally, BSM particles emitted from supernovae can decay to neutrinos en route to Earth, even stronger limits can be set, as shown for instance in \cite{Fiorillo:2022cdq} for Majoron-like bosons and in \cite{Brdar:2023tmi} for a realization featuring neutrino magnetic moment portal. Specifically, unlike neutrinos inside the supernova core, produced BSM states can stream out without further interactions. Consequently, they exit with higher energies than neutrinos, typically around $\mathcal{O}(100)$ MeV. Provided such particles decay to neutrinos, strong limits can be set from the fact that $\mathcal{O}(100)$ MeV energy neutrinos were not recorded from SN 1987A \cite{Fiorillo:2022cdq, Brdar:2023tmi}. So far, such strategies were utilized for neutrino experiments operating during the SN 1987A event as well as future DUNE \cite{DUNE:2015lol} and Hyper-Kamiokande \cite{Hyper-Kamiokande:2018ofw} in light of anticipated galactic supernova event. 
%However, as being decoupled immediately, it is possible that new states can exit the supernova at an earlier time instead---an effect that has not been previously considered.
%This early-glow signal can also be induced by light new states produced from $\nu_e$ scattering inside the supernova given that $\nu_e$ burst happens at an earlier time than the other neutrino flavors \cite{Brdar_2018}.

On the other hand, neutrino telescopes such as IceCube has pioneered in detecting neutrinos at $\mathcal{O}$(PeV) energies \cite{IceCube:2014stg,IceCube:2018cha} and has identified several specific astrophysical neutrino sources \cite{IceCube:2018cha,IceCube:2023ame,IceCube:2022der}.
IceCube is also expected to be able to detect next galactic supernova  \cite{Kopke_2011}. In fact, search for MeV neutrinos from optically obscured galactic supernovae was recently performed \cite{IceCube:2023ogt} as well as the search for temporal correlation of MeV neutrino events with fast radio bursts \cite{IceCube:2019acm}. In both cases, the observable is a collective rise in all photomultiplier rates on top of the background noise  within a certain time window and it turns out that IceCube is sensitive to intervals as short as $\mathcal{O}(0.01)$ s. 
In this work, we demonstrate that such precise time resolution will play a crucial role in testing new physics from supernovae.

 Once produced, the new state can be non-relativistic and travel slowly, causing time delays of the secondary neutrinos. However, it is also possible that new states are rather light implying that they may even exit the supernova at earlier times with respect to certain flavors of SM neutrinos. Such an early BSM signal, which to the best of our knowledge has not been previously considered, can for instance be induced by light BSM states produced from $\nu_e$ scattering inside the supernova; then subsequently decaying to $\bar{\nu}_e$ which chiefly induces the signal at IceCube. This would happen at early stages of supernova, around the neutralization burst, when $\nu_e$ are present and antineutrinos did not get produced yet. Such novel timing patterns generically exist in models with new weakly interacting particles, such as the aforementioned Majoron model \cite{Fiorillo:2022cdq} and neutrino magnetic moment portal \cite{Magill:2018jla,Brdar:2020quo}.
%, but also in scenarios with the mixing portals \cite{Arg_elles_2019,Suliga_2020,Chauhan:2023sci}.
Via the timing measurement, we will show that neutrino telescopes will be able to provide powerful constraints on the BSM parameter space in association to forthcoming galactic supernova event.\\

### 2024-02-21 paper/sections/model.tex
Updates from Overleaf

\textbf{\textit{The Models}}---
As an example, we will consider the Majoron model. Nevertheless the study can also be applied to sterile neutrinos via magnetic portal, as we discuss in the Supplementary Materials. For the case of sterile neutrino with mixing portal, special care needs to be taken for resonant conversions in which case active neutrinos with plasma mass would travel as slowly as the sterile neutrinos at specific distances.
The relevant part of the Lagrangian in the Majoron case reads
\begin{align}
\mathcal{L} \supset -\frac{g_{\alpha\beta}}{2} \nu_\alpha \nu_\beta \phi - m_\phi \phi\phi^* + \text{h.c.}\,,
\end{align}
where $m_\phi$ is the Majoron mass and $g$ parametrizes interaction strength between $\phi$ and neutrinos. Consider flavor universal interaction, we abbreviate $g_{\alpha\beta}\equiv g$. Majorons are produced from neutrino coalescence in the star and then subsequently decay to a pair of neutrinos with the total decay width to all flavor of neutrinos given by $\Gamma_\phi = 3g^2 m_\phi/16 \pi$, and as a result, BSM neutrino flux is produced.

Using data simulated by the Garching group for an $8.8 M_\odot$ progenitor star \cite{Huedepohl2010}, we obtained the fluxes of standard neutrino case that can be fitted to the parametrization in \cite{baxter2021snewpy} and the fluxes of Majoron $\phi$. Requiring the energy loss to Majorons to be less than 10\% of the standard neutrino energy, we can produce the cooling limit, which is in agreement with those obtained in literature \cite{Fiorillo:2022cdq}.
As mentioned above, for the regions below the coupling strength, Majorons produced will immediately stream out. Along its way towards the Earth, Majorons will decay to active neutrinos. For a given emission angle $\alpha$ of daughter neutrino relative to the direction of mother particle, and the angle $\theta$ at which the daughter neutrino arrives at the Earth, the time delay $\Delta t$, the arrival time of the daughter $\nu$ relative to that of the neutrinos produced in the explosion via SM processes, is \cite{Jaeckel:2017tud}
\begin{align}
    \Delta t = L_1/\beta + D_{\rm SN} \cos\theta-L_1 \cos\alpha -D_{\rm SN}\,,
    \label{eq:deltat}
\end{align}
with $D_{\rm SN}$ the distance between the SN and the Earth, $L_1$ the distance mother particles propagated before decaying, and $\beta$ being the speed of $\phi$. The flux of daughter $\nu$ is obtained by considering decays that only occur at $R^{\nu}_{\rm SN}\leq L_1 \leq L^{\rm max}_1$. As the smallest decay distance beyond which the daughter neutrino can escape the explosion unperturbed, we take $R^{\nu}_{\rm SN} = \unit[30]{km}$.
$L^{\rm max}_1$ corresponds to the distance $L_1$ with the largest time delay $\delta t$ \cite{Brdar:2023tmi}.

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_fluxes.pdf}
    \caption{\textit{\textbf{Flux of $\bar{\nu}_{e}$ from Standard Model production and two majoron hypotheses}}.
    We choose the parameters $g_\phi = 10^{-10.2}$ and $g_\phi = 10^{-11.8}$ for $m_\phi = \unit[1]{MeV}$ and $m_\phi = \unit[200]{MeV}$, respectively.
    In the case of the lower-mass model, the majoron can still travel relativistically, leading to a peak in the spectrum at an earlier time.
    In the case of the higher-mass model, the majoron travels subliminally, and thus, the flux peaks at later times.
    Since the dominant interaction is inverse beta decay at these energies, we expect the observed timing distribution to follow the $\bar{\nu}_{e}$
    }
    \label{fig:fluxes}
\end{figure}

For $D_{\rm SN} = \unit[10]{kpc}$ and $\delta t = \unit[100]{s}$ while larger values of $\delta t$ do not affect the limits obtained, we show in \cref{fig:fluxes} the differential fluxes of daughter neutrinos at different time for two benchmarks of the Majoron case, in comparison to the standard neutrino flux.
We can observe that the energy of neutrinos from Majoron decay extends to $\unit[100]{MeV}$ and above, as expected. For heavy Majoron case, the resulting neutrino flux typically is larger at later time, manifesting the time delay from slowly moving Majorons. 
The flux for light Majoron case contain very rich physics. Majorons, generating neutrinos with $E_\nu \geq \unit[150]{MeV}$, are nearly-relativistic with its daughter particles moving in parallel. In such a case, for couplings comparable to or larger than the one shown in \cref{fig:fluxes}, $\Delta t \sim L_d (1/\beta -1)\lesssim \unit[0.01]{sec}$, with $L_d = E_\phi/m_\phi \Gamma^{-1}_\phi \beta$ being the decay length of $\phi$ for the given $\phi$ energy $E_\phi$. The time dependence of neutrino flux thus are mainly inherited from the Majorons upon production, which exhibits in \cref{fig:fluxes} larger fluxes from producing Majorons from the denser and hotter core of SN at earlier time.
For neutrinos with lower energy, i.e $E_\nu \lesssim \unit[40]{MeV}$, the mother Majorons can be non-relativistic, resulting in $\Delta t \gtrsim \unit[0.04]{sec}$. This non-negligible time delay would shift the neutrino flux to later time, lowering the fluxes at earlier time as shown in \cref{fig:fluxes}. For the heavy mass benchmark in \cref{fig:fluxes}, we have $\Delta t \gtrsim \unit[10]{sec}$, which would yield the active neutrino flux at sizeable $t$.

%in details in the following, we consider energy deposits at the photomultipliers, 
%due to denser and hotter core of SN  as Majorans shortly after the bouncing time $t_{\rm bounce}$ from the hot dense core of SN.

>>>>

\textbf{\textit{The Models}}---
%Nevertheless the study can also be applied to sterile neutrinos via magnetic portal, as we discuss in the Supplementary Materials. For the case of sterile neutrino with mixing portal, special care needs to be taken for resonant conversions in which case active neutrinos with plasma mass would travel as slowly as the sterile neutrinos at specific distances.
Let us consider the Majoron model, studied in \cite{Fiorillo:2022cdq}, where supernova 1987A constraints were derived. The relevant part of the Lagrangian reads
\begin{align}
\mathcal{L} \supset -\frac{g_{\alpha\beta}}{2} \nu_\alpha \nu_\beta \phi - m_\phi \phi\phi^* + \text{h.c.}\,,
\end{align}
where $\phi$ is the (pseudo)scalar, $m_\phi$ is the respective mass and $g$ parametrizes interaction strength between $\phi$ and neutrinos. We assume flavor universal interaction and hence $g_{\alpha\beta}\equiv g_\phi \, \delta_{\alpha\beta}$. Majorons are produced from (anti)neutrino coalescence in the star and subsequently decay to a pair of (anti)neutrinos resulting in BSM (anti)neutrino flux emitted from the star. The total decay width is given by $\Gamma_\phi = 3g^2 m_\phi/16 \pi$.

We will set SN bounce time $t_{\rm bounce}$ as $t = 0$. Using data from the simulation of $8.8 M_\odot$ progenitor star \cite{Huedepohl2010, Garching} and including effects of neutrino oscillations assuming normal mass ordering, we calculated standard neutrino fluxes at Earth that can be presented in terms of the parametrization with the ``pinching'' factor \cite{Keil:2002in,baxter2021snewpy}. Following \cite{Fiorillo:2022cdq}, we also calculated the flux of emitted Majorons. At $t \leq \unit[0.05]{s}$, Majorons are mainly produced through 
$\nu_e$ and $\nu_x$ coalescence with $\nu_x$ including (anti)neutrinos of $\tau$ and $\mu$ flavor. 
After $\unit[0.05]{s}$, the flux of Majorons decreases as the total flux of neutrinos of all flavors. Thus we found that Majoron flux peaks around $t \sim \unit[0.05]{s}$.
We further cross checked the agreement with \cite{Fiorillo:2022cdq} by reproducing the bound from SN energy loss criteria.

Very weakly coupled Majorons immediately stream out after the production. On their way to Earth, Majorons of energy $E_\phi$ will travel at the speed of $\beta$ for a distance of $L_1$, then decay to (anti)neutrinos of energy $E_\nu$ at an emission angle $\cos\alpha = (2 E_\phi E_{\nu} - M^2_N)/(2E_{\nu}E_\phi\beta)$. Daughter (anti)neutrino will travel for a distance $L_2$ and reach the Earth at the angle $\theta$ satisfying $D_{\rm SN} \sin\theta = L_1 \sin\alpha$ for a SN that is $D_{\rm SN}$ away (see such geometry in \cite{Brdar:2023tmi}). The time delay, $\delta t$, relative to a relativistic particle moving along the line-of-sight path from SN to reach the Earth is given by
\cite{Jaeckel:2017tud}
\begin{align}
    \delta t = (\frac{L_1}{\beta}-L_1) + (L_1 + L_2 -D_{\rm SN})\,,
    \label{eq:deltatexa}
\end{align}
where we have split the contribution to $\delta t$ into two parts: the first part comes from the slowly moving of Majorons before its decay, and the second part is due to the detour from the line-of-sight path at the given emission angle $\alpha$.

To get the characteristic value of $\delta t$, we can take $L_1 = L_\phi$ with $L_\phi = (E_\phi/m_\phi) \Gamma^{-1}_\phi \beta$ being the decay length of $\phi$. As for the parameter regions we considered, $(D_{\rm SN} -D_{\rm SN} \cos\theta) \ll \unit[10^{-3}]{s}$, $\delta t$ can be approximated
\begin{align}
    \delta t &\simeq L_\phi(\frac{1}{\beta} -1) + L_\phi(1-\cos\alpha) = \frac{8\pi}{3 E_{\bar{\nu}_{e}}g_\phi^2}\,,
    \label{eq:deltat}
\end{align}
which is roughly independent of Majoron energy $E_\phi$. However, at small values of $E_\phi$, the main contribution to $\delta t$ comes from the first term due to slowing moving Majorons, while at larger values of $E_\phi$, from the second term due to larger detour given the increase of $L_\phi$ over the decrease of $(1-\cos\alpha)$.
%For neutrinos from light Majoron decay, the characeristic time delay is set by the detour, while for   
%We have $L_1 \lesssim L_\phi$ with $L_\phi = (E_\phi/m_\phi) \Gamma^{-1}_\phi \beta$ being the decay length of $\phi$.
%As $L_\phi$ is much smaller than $D_{\rm SN}$ for all values of $E_\phi$ and the parameter regions considered, we can estimate that $(D_{\rm SN} -D_{\rm SN} \cos\theta) \ll \unit[10^{-3}]{s}$. Thus we can approximate $\Delta t$ as
%\begin{align}
%    \delta t &\simeq L_1/\beta -L_1 \cos\alpha\ \\\notag
%    &= (L_1/\beta -L_1) + (L_1-L_1 \cos\alpha)\,.
%    \label{eq:deltat}
%\end{align}
%Taking $L_1 \simeq L_\phi$, we have $\Delta t \sim 8\pi/(3 E_e g_\phi^2)$.

We can obtain the flux of daughter $\nu$ by considering decays that occur only at $R^{\nu}_{\rm SN}\leq L_1$. As the smallest decay distance beyond which the daughter neutrino can escape the explosion unperturbed, we take $R^{\nu}_{\rm SN} = \unit[30]{km}$. We will also limit our analysis to a maximal time window of $\unit[100]{s}$ after SN bounce. This time window is larger than typical values of time delay considered for the parameter space in consideration.
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_fluxes.pdf}
    \caption{\textit{\textbf{Flux of $\bar{\nu}_{e}$ from Standard Model production and two majoron hypotheses}}.
    We choose the parameters $g_\phi = 10^{-10.2}$ and $g_\phi = 10^{-11.8}$ for $m_\phi = \unit[1]{MeV}$ and $m_\phi = \unit[200]{MeV}$, respectively. \yy{the unit for the x-axis should be MeV, for y-axis, we chose to use $\rm s$ for second instead of $\rm sec$.}
    }
    \label{fig:fluxes}
\end{figure}
We will consider a SN event that happens in the galaxy at a distance $D_{\rm SN}=\unit[10]{kpc}$, which is not unlikely~\cite{Reed:2005en,Rozwadowska:2020nab}. As $\bar{\nu}_e$ contribute chiefly to the signals at IceCube via inverse beta decay (IBD), we show in \cref{fig:fluxes} the differential fluxes of daughter $\bar{\nu}_e$ at different time for two benchmark points in the Majoron model, in comparison to the SM $\bar{\nu}_e$ flux. We expect the observed signal timing distribution at IceCube to follow these patterns of $\bar{\nu}_{e}$ flux.

From \cref{fig:fluxes}, we can observe that the energy of $\bar{\nu}_e$ from Majoron decay extends to $\unit[100]{MeV}$ and above for both benchmarks, as expected. 
Consider the light Majoron case in \cref{fig:fluxes}. Majorons, generating neutrinos with $E_{\bar{\nu}_e} \geq \unit[150]{MeV}$, are nearly-relativistic with the emission angle $\alpha$ close to zero. The time delay can be estimated from the first term in \cref{eq:deltat} given by $\delta t \sim L_\phi (1/\beta -1)\lesssim \unit[0.01]{s}$. With such a negligible time delay, the time dependence of the $\bar{\nu}_e$ flux is mainly inherited from that of the Majorons upon production, with larger fluxes at $t\leq \unit[0.05]{s}$. Notice that such $\bar{\nu}_e$ flux would arrive at the detector earlier than the peak of the SM $\bar{\nu}_e$ flux which is around $t\simeq \unit[0.1]{s}$ from \cref{fig:fluxes}, it potentially leads to early signals at IceCube.
%For $\bar{\nu}_e$ produced with energy, i.e $E_{\bar{\nu}_{e}} \sim \unit[10]{MeV}$, a fraction of their mother Majorons will have energy $E_\phi\sim\mathcal{O}(\unit[10]{MeV})$, resulting in typical $\delta t \gtrsim \unit[0.1]{s}$. For higher values of $E_\phi$, the main contribution from detour also leads to $\delta t \gtrsim \unit[0.1]{s}$. 
The time delay for $\bar{\nu}_e$ with energy, i.e $E_{\bar{\nu}_{e}} \sim \unit[10]{MeV}$ is typically $\gtrsim \unit[0.1]{s}$ from \cref{eq:deltat}. These $\bar{\nu}_e$ could be produced from Majorons with low energy of $\mathcal{O}(\unit[10]{MeV})$ or high energies. For the low energy Majoron case, the time delay is mainly due to slowly moving Majoron, while for high energy case, mainly from the detour.
Such sizable time delay will shift $\bar{\nu}_e$ peak flux to later time, in comparison to that of their mother Majorons, as exhibited in \cref{fig:fluxes}.

For heavy Majoron case, the resulting $\bar{\nu}_e$ flux is larger at later time, manifesting the time delay from slowly moving Majorons. For the heavy Majoron benchmark in \cref{fig:fluxes}, we can estimate that $\delta t \gtrsim \unit[10]{s}$ for $E_{\bar{\nu}_{e}}\sim \unit[200]{MeV}$ which is delayed comparing to the peak of SM $\bar{\nu}_e$ flux. Comparing to the light Majoron case, such large time delay is a combination of a more slowly moving Majoron and a larger detour due to its large emission angle. As $L_1\lesssim L_\phi$, the time spread of the flux is roughly given by the aforementioned characteristic value of time delay which is $\unit[0.1]{s}$ ($\gtrsim\unit[10]{s}$) for the light (heavy) benchmark case.
%which yield sizable $\bar{\nu}_e$ flux extending to sizeable $t$.
%in details in the following, we consider energy deposits at the photomultipliers, 
%due to denser and hotter core of SN  as Majorans shortly after the bouncing time $t_{\rm bounce}$ from the hot dense core of SN.
